---
title: "Assignment 2"
author: "Nathan Weitershausen"
format: html
editor: visual
embed-resources: true
---

```{r}
library(dplyr)
library(ggplot2)
library(nycflights13)
airlines <- airlines
airports <- airports
flights <- flights
planes <- planes
weather <- weather
```

# Question 1

Find the top 10 most popular destinations in the flights data. How many flights went to each?

```{r}
Top10Flights <- sort(table(flights$dest), decreasing = TRUE)

Top10Flights <- head(Top10Flights, 10)

Top10Flights
```

## Summary

The 10 most popular destinations in the flights data, in order from most to least popular and with the number of flights that went to each in parantheses, are: ORD (17,283), ATL (17,215), LAX (16,174), BOS (15,508), MCO (14082), CLT (14,064), SFO (13,331), FLL (12,055), MIA (11,728), and DCA (9705).

# Question 2

Create two new categorical variables in flights based on the departure time (dep_time), and arrival time (arr_time). These new variables should have four categories: “early morning” for 12am to 6am, “morning” for 6am to 12pm, “afternoon” for 12pm to 6pm, and “evening” for 6pm to 12am. Show barplots of both of these new variables. What percentage of flights were “red eye” flights? We’ll define these as flights that depart in “afternoon” or “evening” and arrive in “early morning” or “morning.”

```{r}
flights$dep_time_category <- ifelse(is.na(flights$dep_time), NA, ifelse(flights$dep_time < 600, "Early Morning", ifelse(flights$dep_time < 1200, "Morning", ifelse(flights$dep_time < 1800, "Afternoon", "Evening"))))

flights$arr_time_category <- ifelse(is.na(flights$arr_time), NA, ifelse(flights$arr_time < 600, "Early Morning", ifelse(flights$arr_time < 1200, "Morning", ifelse(flights$arr_time < 1800, "Afternoon", "Evening"))))

Time_Labels <- c("Early Morning", "Morning", "Afternoon", "Evening")

barplot(table(flights$dep_time_category), main = "Flights by Departure Time Category", xlab = "Departure Time", ylab = "Number of Flights", ylim = c(0, 160000))

barplot(table(flights$arr_time_category), main = "Flights by Arrival Time Category", xlab = "Arrival Time", ylab = "Number of Flights", ylim = c(0, 160000))
```
```{r}
# Percentage of Flights that were Red Eye Flights (Depart: Afternoon and Evening, Arrive: Early Morning and Morning)

flights <- flights[!is.na(flights$dep_time) & !is.na(flights$arr_time),]
flights$red_eye <- flights$dep_time_category %in% c("Afternoon", "Evening") & flights$arr_time_category %in% c("Early Morning", "Morning")

RedEyePercentage <- mean(flights$red_eye, na.rm = TRUE) * 100

RedEyePercentage

RedEyeNumber <- table(flights$red_eye)

RedEyePercentage2 <- data.frame(Red_Eye = names(RedEyeNumber), Amount = as.vector(RedEyeNumber), Percentage = as.vector(prop.table(RedEyeNumber)*100))

RedEyePercentage2
```

## Summary

About 3.24% of flights were Red Eye Flights. 

# Question 3

Were there any planes that flew for multiple airlines? If so, how many were there and which airlines did they fly for?

```{r}
Plane_Airlines_Carriers <- unique(flights[c("tailnum", "carrier")])
Carriers_by_Tailnum <- split(Plane_Airlines_Carriers$carrier, Plane_Airlines_Carriers$tailnum)
Carriers_by_Tailnum <- Carriers_by_Tailnum[sapply(Carriers_by_Tailnum, length) >1]
as.data.frame(Carriers_by_Tailnum)
```

## Summary

There were 17 planes that flew for multiple airlines. Planes N146PQ, N153PQ, N176PQ, N181PQ, N197PQ, N200PQ, N228PQ, and N232PQ flew for airlines 9E and EV. Planes N933AT, N935AT, N977AT, N978AT, N979AT, N981AT, N989AT, N990AT, and N994AT flew for airlines FL and DL. 

# Question 4

In the figure above, there is a missing relationship between weather and airports. What is the relationship and how should it appear in the diagram?

```{r}
weather
airports
```

## Summary

The missing relationship between weather and airports is the airport of origin from which the weather observation was made. In weather, the airport of origin is under the column "origin." In airports, the column "faa" refers to the FAA code of the airport. However, it appears that there are only three origins in weather, those being EWR, LGA, and JFK, while airports has 1,458 FAA codes. In the diagram, a connection should be made between weather "origin" and airports "faa."

# Question 5

## Part 1
In order to address our primary question, we want to prepare a tidy dataset (a single data frame). This will consist of information from the flights and weather data. The year, month, day, hour, and origin variables from weather provide almost enough information to give each observation a unique value. Create a new variable in the weather dataset by pasting together those 5 variables, then determine how many duplicated values there are and explain why (Hint: compare your new variable to the time_hour variable).

```{r}
weather$Unique_Weather <- paste(weather$year, weather$month, weather$day, weather$hour, weather$origin)

sum(duplicated(weather$Unique_Weather))

weather[duplicated(weather$Unique_Weather, fromLast = TRUE), c("time_hour", "Unique_Weather")]
```

### Summary

There are 3 duplicated values in the newly created "Unique_Weather" variable, occurring on November 3rd, 2013 at 1:00 AM at each airport. This is because this day was the end of Daylight Saving time, and clocks were moved back an hour, so two observations of 1:00 AM were recorded at each airport. 

## Part 2
Now merge the flights data and the weather data so that each flight contains information about the weather at its departure airport at the time it was scheduled to take off. You’ll want to use the variables time_hour and origin in order to do this.

```{r}
Merged_Flights_Weather <- merge(flights, weather, by = c("time_hour", "origin"), all.x = TRUE)

Merged_Flights_Weather <- subset(Merged_Flights_Weather, select = - c(year.y, month.y, day.y, hour.y))

names(Merged_Flights_Weather)[names(Merged_Flights_Weather) == "year.x"] <- "year"
names(Merged_Flights_Weather)[names(Merged_Flights_Weather) == "month.x"] <- "month"
names(Merged_Flights_Weather)[names(Merged_Flights_Weather) == "day.x"] <- "day"
names(Merged_Flights_Weather)[names(Merged_Flights_Weather) == "hour.x"] <- "hour"
```

# Question 6

On this merged dataset, perform steps 2-5 of the EDA checklist presented in class. Remember the context provided by our primary question above.

EDA Checklist: 2. Check the size of the data. 3. Examine the variables and their types. 4. Look at the top and bottom of the data. 5. Visualize the distributions of key variables. 

```{r}
# 2. Check the size of the data
dim(Merged_Flights_Weather)
nrow(Merged_Flights_Weather)
ncol(Merged_Flights_Weather)

# 3. Examine the variables and their types
str(Merged_Flights_Weather)

# 4. Look at the top and bottom of the data
head(Merged_Flights_Weather)
tail(Merged_Flights_Weather)

# 5. Visualize the distributions of key variables
barplot(table(Merged_Flights_Weather$origin), main = "Number of Flights by Origin Airport", xlab = "Origin Airport", ylab = "Number of Flights", ylim = c(0, 140000))
mean(is.na(Merged_Flights_Weather$origin))

hist(Merged_Flights_Weather$dep_delay, main = "Distribution of Departure Delays", xlab = "Departure Delay (minutes)", sub = "Negative Values Represent Earlier Depature")
summary(Merged_Flights_Weather$dep_delay)
mean(is.na(Merged_Flights_Weather$dep_delay))

hist(Merged_Flights_Weather$arr_delay, main = "Distribution of Arrival Delays", xlab = "Arrival Delay (minutes)", sub = "Negative Values Represent Earlier Arrival")
summary(Merged_Flights_Weather$arr_delay)
mean(is.na(Merged_Flights_Weather$arr_delay))

hist(Merged_Flights_Weather$temp, main = "Distribution of Temperatures at Origin Airports", xlab = "Temperature (Fahrenheit)")
summary(Merged_Flights_Weather$temp)
mean(is.na(Merged_Flights_Weather$temp))

hist(Merged_Flights_Weather$dewp, main = "Distribution of Dewpoint at Origin Airports", xlab = "Dew Point (Fahrenheit)")
summary(Merged_Flights_Weather$dewp)
mean(is.na(Merged_Flights_Weather$dewp))

hist(Merged_Flights_Weather$humid, main = "Distribution of Humidity at Origin Airports", xlab = 
"Humidity")
summary(Merged_Flights_Weather$humid)
mean(is.na(Merged_Flights_Weather$humid))

hist(Merged_Flights_Weather$wind_speed, main = "Distribution of Wind Speed at Origin Airports", xlab = "Wind Speed (mph)")
summary(Merged_Flights_Weather$wind_speed)
mean(is.na(Merged_Flights_Weather$wind_speed))

hist(Merged_Flights_Weather$wind_gust, main = "Distribution of Wind Gust at Origin Airports", xlab = "Wind Gust (mph)")
summary(Merged_Flights_Weather$wind_gust)
mean(is.na(Merged_Flights_Weather$wind_gust))

hist(Merged_Flights_Weather$precip, main = "Distribution of Precipitation at Origin Airports", xlab = "Precipitation (inches)")
summary(Merged_Flights_Weather$precip)
mean(is.na(Merged_Flights_Weather$precip))

hist(Merged_Flights_Weather$pressure, main = "Distribution of Atmospheric Pressure at Origin Airports", xlab = "Atmospheric Pressure (millibars)")
summary(Merged_Flights_Weather$pressure)
mean(is.na(Merged_Flights_Weather$pressure))

hist(Merged_Flights_Weather$visib, main = "Distribution of Visibility at Origin Airports", xlab = "Visibility (miles)")
summary(Merged_Flights_Weather$visib)
mean(is.na(Merged_Flights_Weather$visib))
```

## Summary
There are 328,063 recordings (rows) and 32 variables (columns). The variables are mainly integer and numeric, but there is one logical variable and some character variables. In the context of the primary question, the key variables would be: dep_delay, arr_delay, temp, dewp, humid, wind_speed, wind_gust, precip, pressure, and visib. Measurements units were found using the data library for flights and weather data. The data was not cleaned in this step, so there are still some missing values. 

# Question 7

## Part 1

Calculate the average departure delay for each day (Hint: think about which variables you need in order to do this!). Which day had the worst average length of delay for departures? 

```{r}
Daily_Delay <- Merged_Flights_Weather |>
  group_by(year, month, day) |>
  summarize(Average_Departure_Delay = mean(dep_delay, na.rm = TRUE), Amount_of_Flights = n()) |>
  arrange(desc(Average_Departure_Delay))

head(Daily_Delay)
```

### Summary

March 8th, 2013 was the worst day, as it had an average departure delay of ~84 minutes. 

## Part 2

Now calculate the averages by day and origin. Which airport had the worst single day for delays and when was it? 

```{r}
Daily_Delay_Airport <- Merged_Flights_Weather |>
  group_by(year, month, day, origin) |>
  summarize (Average_Depatrue_Delay = mean(dep_delay, na.rm = TRUE), Amount_of_Flights = n()) |>
  arrange(desc(Average_Depatrue_Delay))

head(Daily_Delay_Airport)
```


### Summary

LGA had the single worst day for delays on March 8th, 2013, with an average departure delay of ~106 minutes. 

## Part 3

Now calculate the averages by hour and origin. Which airport had the worst single hour for delays (and when was it)?

```{r}
Hourly_Delay_Airport <- Merged_Flights_Weather |>
  group_by(time_hour, origin) |>
  summarize(Average_Depature_Delay = mean(dep_delay, na.rm = TRUE), Amount_of_Flights = n()) |>
  arrange(desc(Average_Depature_Delay))

head(Hourly_Delay_Airport)
```


### Summary

LGA had the worst single hour for delays on July 22nd, 2013 at 6 PM, with an average departure delay of ~310 minutes.

# Question 8

Compute the average arrival delay by destination, then add that information to the airports data frame. Make a map showing the spatial distribution of delays (you can make it however you like, but we recommend using either ggplot or leaflet). Use the color of the points to display the average delay for each airport.

```{r}
Arrival_Delay <- Merged_Flights_Weather |>
  group_by(dest) |>
  summarize(Average_Arrival_Delay = mean(arr_delay, na.rm = TRUE), Amount_of_Flights = sum(!is.na(arr_delay))) |>
  filter(Amount_of_Flights > 0)

airports <- airports |>
  left_join(Arrival_Delay, by = c("faa" = "dest"))

airports <- airports |>
  filter(!is.na(Average_Arrival_Delay))

US_Map <- map_data("state")
ggplot() +
  geom_polygon(data = US_Map, aes(x = long, y = lat, group = group), fill = "gray", color = "white") +
  geom_point(data=airports,  aes(x = lon, y = lat, color = Average_Arrival_Delay)) + 
  scale_color_gradient2(low = "darkblue", mid = "orange", high = "cyan", name = "Average Arrival Delay (Minutes)") +
  labs(title = "Average Arrival Delay by Arrival Airport", x = "Longitude", y = "Latitude")
```

# Question 9

Finally, create as many plots and summary statistics as needed to answer the primary question: Which weather phenomena have the most impact on flight delays?

```{r}
Weather_Variables <- c("dep_delay", "temp", "humid", "dewp", "wind_speed", "wind_gust", "precip", "pressure", "visib")

Weather_Delay <- Merged_Flights_Weather
Weather_Delay$Visibility_Category <- ifelse(is.na(Weather_Delay$visib), NA, ifelse(Weather_Delay$visib < 3, "Low Visibility", "High Visibility"))
Weather_Delay$Precipitation_Category <- ifelse(is.na(Weather_Delay$precip), NA, ifelse(Weather_Delay$precip > 0, "Precipitation", "No Precipitation"))
Weather_Delay$WindSpeed_Category <- ifelse(is.na(Weather_Delay$wind_speed), NA, ifelse(Weather_Delay$wind_speed >= 30, "High Wind", "Low Wind"))
Weather_Delay$WindGust_Category <- ifelse(is.na(Weather_Delay$wind_gust), NA, ifelse(Weather_Delay$wind_gust >= 30, "High Gust", "Low Gust"))
Weather_Delay$Pressure_Category <- ifelse(is.na(Weather_Delay$pressure), NA, ifelse(Weather_Delay$pressure >= 1015.92, "High Pressure", "Low Pressure"))
Weather_Delay$Dewpoint_Category <- ifelse(is.na(Weather_Delay$dewp), NA, ifelse(Weather_Delay$dewp >= 65, "High Dew Point", "Low Dew Point"))
Weather_Delay$Temperature_Category <- ifelse(is.na(Weather_Delay$temp), NA, ifelse(Weather_Delay$temp >= 100, "High Temp", "Low Temp"))
Weather_Delay$Humidity_Category <- ifelse(is.na(Weather_Delay$humid), NA, ifelse(Weather_Delay$humid >= 60, "High Humidity", "Low Humidity"))

Correlations <- cor(Weather_Delay[Weather_Variables], use = "pairwise.complete.obs", method = "spearman")

Departure_Delay_Correlation <- Correlations[rownames(Correlations) != "dep_delay", "dep_delay"]
Departure_Delay_Correlation <- Departure_Delay_Correlation[order(Departure_Delay_Correlation, decreasing = TRUE)]
Departure_Delay_Correlation

barplot(Departure_Delay_Correlation, horiz = TRUE, las = 1, col = ifelse(Departure_Delay_Correlation >0, "red", "blue"), xlab = "Correlation with Depature Delay", main = "Weather and Depature Delays")
```

```{r}
High_Visibility_Mean <- mean(Weather_Delay$dep_delay[Weather_Delay$Visibility_Category == "High Visibility"], na.rm = TRUE)

Low_Visibitiliy_Mean <- mean(Weather_Delay$dep_delay[Weather_Delay$Visibility_Category == "Low Visibility"], na.rm = TRUE)

Visibility_Delay <- data.frame(Visibility_Category = c("High Visibility", "Low Visibility"), Departure_Delay_Mean = c(High_Visibility_Mean, Low_Visibitiliy_Mean))
Visibility_Delay

barplot(Visibility_Delay$Departure_Delay_Mean, names.arg = Visibility_Delay$Visibility_Category, col = c("red", "blue"), ylab = "Average Departure Delay (minutes)", xlab = "Visbility Category", main = "Average Depature Delay by Visibility", sub = "A higher average for Low Visibility indicates a longer depature delay")

plot(Weather_Delay$visib, Weather_Delay$dep_delay, xlab = "Visibility (miles)", ylab = "Departure Delay (minutes)", main = "Departure Delay by Visbility", xlim = c(0,10))
```

```{r}
No_Precipitation_Mean <- mean(Weather_Delay$dep_delay[Weather_Delay$Precipitation_Category == "No Precipitation"], na.rm = TRUE)

Precipitation_Mean <- mean(Weather_Delay$dep_delay[Weather_Delay$Precipitation_Category == "Precipitation"], na.rm = TRUE)

Precipitation_Delay <- data.frame(Precipitation_Category = c("No Precipitation", "Precipitation"), Departure_Delay_Mean = c(No_Precipitation_Mean, Precipitation_Mean))
Precipitation_Delay

barplot(Precipitation_Delay$Departure_Delay_Mean, names.arg = Precipitation_Delay$Precipitation_Category, col = c("red", "blue"), ylab = "Average Departure Delay (minutes)", xlab = "Precipitation Category", main = "Average Depature Delay by Precipitation Status", sub = "A higher average for Precipitation indicates a longer depature delay")

plot(Weather_Delay$precip, Weather_Delay$dep_delay, xlab = "Precipitaton (inches)", ylab = "Departure Delay (minutes)", main = "Departure Delay by Precipitation")
```

```{r}
High_Wind_Mean <- mean(Weather_Delay$dep_delay[Weather_Delay$WindSpeed_Category == "High Wind"], na.rm = TRUE)

Low_Wind_Mean <- mean(Weather_Delay$dep_delay[Weather_Delay$WindSpeed_Category == "Low Wind"], na.rm = TRUE)

WindSpeed_Delay <- data.frame(WindSpeed_Category = c("High Wind", "Low Wind"), Departure_Delay_Mean = c(High_Wind_Mean, Low_Wind_Mean))
WindSpeed_Delay

barplot(WindSpeed_Delay$Departure_Delay_Mean, names.arg = WindSpeed_Delay$WindSpeed_Category, col = c("red", "blue"), ylab = "Average Departure Delay (minutes)", xlab = "Wind Speed Category", main = "Average Depature Delay by Wind Speed", sub = "A higher average for High Wind Speed indicates a longer depature delay")

plot(Weather_Delay$wind_speed, Weather_Delay$dep_delay, xlab = "Wind Speed (mph)", ylab = "Departure Delay (minutes)", main = "Departure Delay by Wind Speed")

High_Gust_Mean <- mean(Weather_Delay$dep_delay[Weather_Delay$WindGust_Category == "High Gust"], na.rm = TRUE)

Low_Gust_Mean <- mean(Weather_Delay$dep_delay[Weather_Delay$WindGust_Category == "Low Gust"], na.rm = TRUE)

WindGust_Delay <- data.frame(WindGust_Category = c("High Gust", "Low Gust"), Departure_Delay_Mean = c(High_Gust_Mean, Low_Gust_Mean))
WindGust_Delay

barplot(WindGust_Delay$Departure_Delay_Mean, names.arg = WindGust_Delay$WindGust_Category, col = c("red", "blue"), ylab = "Average Departure Delay (minutes)", xlab = "Wind Gust Category", main = "Average Depature Delay by Wind Gust", sub = "A higher average for High Wind Gust indicates a longer depature delay")

plot(Weather_Delay$wind_gust, Weather_Delay$dep_delay, xlab = "Wind Gust (mph)", ylab = "Departure Delay (minutes)", main = "Departure Delay by Wind Gust")
```

```{r}
High_Pressure_Mean <- mean(Weather_Delay$dep_delay[Weather_Delay$Pressure_Category == "High Pressure"], na.rm = TRUE)

Low_Pressure_Mean <- mean(Weather_Delay$dep_delay[Weather_Delay$Pressure_Category == "Low Pressure"], na.rm = TRUE)

Pressure_Delay <- data.frame(Pressure_Category = c("High Pressure", "Low Pressure"), Departure_Delay_Mean = c(High_Pressure_Mean, Low_Pressure_Mean))
Pressure_Delay

barplot(Pressure_Delay$Departure_Delay_Mean, names.arg = Pressure_Delay$Pressure_Category, col = c("red", "blue"), ylab = "Average Departure Delay (minutes)", xlab = "Atmospheric Pressure Category", main = "Average Depature Delay by Pressure Status", sub = "A higher average for Low Pressure indicates a longer depature delay")

plot(Weather_Delay$pressure, Weather_Delay$dep_delay, xlab = "Pressure (millibars)", ylab = "Departure Delay (minutes)", main = "Departure Delay by Atmospheric Pressure")
```

```{r}
High_Dewp_Mean <- mean(Weather_Delay$dep_delay[Weather_Delay$Dewpoint_Category == "High Dew Point"], na.rm = TRUE)

Low_Dewp_Mean <- mean(Weather_Delay$dep_delay[Weather_Delay$Dewpoint_Category == "Low Dew Point"], na.rm = TRUE)

Dewp_Delay <- data.frame(Dewpoint_Category = c("High Dew Point", "Low Dew Point"), Departure_Delay_Mean = c(High_Dewp_Mean, Low_Dewp_Mean))
Dewp_Delay

barplot(Dewp_Delay$Departure_Delay_Mean, names.arg = Dewp_Delay$Dewpoint_Category, col = c("red", "blue"), ylab = "Average Departure Delay (minutes)", xlab = "Dew Point Category", main = "Average Depature Delay by Dew Point", sub = "A higher average for High Dew Point indicates a longer depature delay")

plot(Weather_Delay$dewp, Weather_Delay$dep_delay, xlab = "Dew Point (Fahrenheit)", ylab = "Departure Delay (minutes)", main = "Departure Delay by Dew Point")

High_Temp_Mean <- mean(Weather_Delay$dep_delay[Weather_Delay$Temperature_Category == "High Temp"], na.rm = TRUE)

Low_Temp_Mean <- mean(Weather_Delay$dep_delay[Weather_Delay$Temperature_Category == "Low Temp"], na.rm = TRUE)

Temp_Delay <- data.frame(Temperature_Category = c("High Temperature", "Low Temperature"), Departure_Delay_Mean = c(High_Temp_Mean, Low_Temp_Mean))
Temp_Delay

barplot(Temp_Delay$Departure_Delay_Mean, names.arg = Temp_Delay$Temperature_Category, col = c("red", "blue"), ylab = "Average Departure Delay (minutes)", xlab = "Temperature Category", main = "Average Depature Delay by Temperature", sub = "A higher average for High Temperature indicates a longer depature delay")

plot(Weather_Delay$temp, Weather_Delay$dep_delay, xlab = "Temperature (Fahrenheit)", ylab = "Departure Delay (minutes)", main = "Departure Delay by Temperature")

High_Humid_Mean <- mean(Weather_Delay$dep_delay[Weather_Delay$Humidity_Category == "High Humidity"], na.rm = TRUE)

Low_Humid_Mean <- mean(Weather_Delay$dep_delay[Weather_Delay$Humidity_Category == "Low Humidity"], na.rm = TRUE)

Humid_Delay <- data.frame(Humidity_Category = c("High Humidity", "Low Humidity"), Departure_Delay_Mean = c(High_Humid_Mean, Low_Humid_Mean))
Humid_Delay

barplot(Humid_Delay$Departure_Delay_Mean, names.arg = Humid_Delay$Humidity_Category, col = c("red", "blue"), ylab = "Average Departure Delay (minutes)", xlab = "Humidity Category", main = "Average Depature Delay by Humidity", sub = "A higher average for High Humidity indicates a longer depature delay")

plot(Weather_Delay$humid, Weather_Delay$dep_delay, xlab = "Humidity", ylab = "Departure Delay (minutes)", main = "Departure Delay by Humidity")

```


## Notes

Low visibility was determined as anything less than 3 miles. High wind and high gust were determined to be anything 30+ mph. High pressure was determined to be anything above 1015.92 millibars (30 inHg). https://www.faa.gov/nextgen/programs/weather/faq was used as a source.

## Summary
Atmospheric pressure and visibility were found to have a weak negative association with departure delay. Though flights experienced longer delays under low visibility and atmospheric pressure conditions, both of which are associated with rain and storms, the correlations suggest that these two variables alone had limited influence on departure delays. Wind gust, temperature, wind speed, dew point, humidity, and precipitation had weak positive associations with departure delay, with precipitation having the strongest association at ~0.12. Flights experienced delays in high wind speed, wind gust, dew point, humidity, and temperature conditions,as well as when it was precipitating. All of these conditions are associated with storms. Taking all of these variables into account, including visibility and atmospheric pressure, it appears that storm conditions is the weather phenomena that has the most impact on flight delays.

```{r}
Visibility <- Merged_Flights_Weather |>
  mutate(Visibility_Category = case_when(is.na(visib) ~ NA_character_, visib < 3 ~ "Low Visibility", visib >= 3 ~ "High Visibility"))

Visibility_Airport <- Visibility |>
  distinct(origin, time_hour, .keep_all = TRUE) |>
  filter(!is.na(Visibility_Category)) |>
  count(origin, Visibility_Category, name = "Number of Hours") |>
  group_by(origin)

Visibility_Airport <- Visibility_Airport |>
  select(-any_of(c("lat","lon"))) |>
  left_join(nycflights13::airports, by= c("origin" = "faa"))

ggplot() +
  geom_polygon(data = US_Map, aes(x = long, y = lat, group = group), fill = "gray", color = "white") +
  geom_point(data = Visibility_Airport, aes(x = lon, y = lat, color = Visibility_Category)) + 
  coord_quickmap(xlim= c(-75, -72), ylim = c(40, 41.5)) +
  scale_color_manual(values = c("Low Visibility" = "Red", "High Visibility" = "Green"))
```

## Summary
