---
title: "Assignment 2"
author: "Nathan Weitershausen"
format: html
editor: visual
embed-resources: true
---

```{r}
library(dplyr)
library(ggplot2)
library(nycflights13)
airlines <- airlines
airports <- airports
flights <- flights
planes <- planes
weather <- weather
```

# Question 1

Find the top 10 most popular destinations in the flights data. How many flights went to each?

```{r}
Top10Flights <- sort(table(flights$dest), decreasing = TRUE)

Top10Flights <- head(Top10Flights, 10)

Top10Flights
```

## Summary

The 10 most popular destinations in the flights data, in order from most to least popular and with the number of flights that went to each in parantheses, are: ORD (17,283), ATL (17,215), LAX (16,174), BOS (15,508), MCO (14082), CLT (14,064), SFO (13,331), FLL (12,055), MIA (11,728), and DCA (9705).

# Question 2

Create two new categorical variables in flights based on the departure time (dep_time), and arrival time (arr_time). These new variables should have four categories: “early morning” for 12am to 6am, “morning” for 6am to 12pm, “afternoon” for 12pm to 6pm, and “evening” for 6pm to 12am. Show barplots of both of these new variables. What percentage of flights were “red eye” flights? We’ll define these as flights that depart in “afternoon” or “evening” and arrive in “early morning” or “morning.”

```{r}
flights$dep_time_category <- ifelse(is.na(flights$dep_time), NA, ifelse(flights$dep_time < 600, "Early Morning", ifelse(flights$dep_time < 1200, "Morning", ifelse(flights$dep_time < 1800, "Afternoon", "Evening"))))

flights$arr_time_category <- ifelse(is.na(flights$arr_time), NA, ifelse(flights$arr_time < 600, "Early Morning", ifelse(flights$arr_time < 1200, "Morning", ifelse(flights$arr_time < 1800, "Afternoon", "Evening"))))

Time_Labels <- c("Early Morning", "Morning", "Afternoon", "Evening")

barplot(table(flights$dep_time_category), main = "Flights by Departure Time Category", xlab = "Departure Time", ylab = "Number of Flights", ylim = c(0, 160000))

barplot(table(flights$arr_time_category), main = "Flights by Arrival Time Category", xlab = "Arrival Time", ylab = "Number of Flights", ylim = c(0, 160000))
```

```{r}
# Percentage of Flights that were Red Eye Flights (Depart: Afternoon and Evening, Arrive: Early Morning and Morning)

flights <- flights[!is.na(flights$dep_time) & !is.na(flights$arr_time),]
flights$red_eye <- flights$dep_time_category %in% c("Afternoon", "Evening") & flights$arr_time_category %in% c("Early Morning", "Morning")

RedEyePercentage <- mean(flights$red_eye, na.rm = TRUE) * 100

RedEyePercentage

RedEyeNumber <- table(flights$red_eye)

RedEyePercentage2 <- data.frame(Red_Eye = names(RedEyeNumber), Amount = as.vector(RedEyeNumber), Percentage = as.vector(prop.table(RedEyeNumber)*100))

RedEyePercentage2
```

## Summary

About 3.24% of flights were Red Eye Flights.

# Question 3

Were there any planes that flew for multiple airlines? If so, how many were there and which airlines did they fly for?

```{r}
Plane_Airlines_Carriers <- unique(flights[c("tailnum", "carrier")])
Carriers_by_Tailnum <- split(Plane_Airlines_Carriers$carrier, Plane_Airlines_Carriers$tailnum)
Carriers_by_Tailnum <- Carriers_by_Tailnum[sapply(Carriers_by_Tailnum, length) >1]
as.data.frame(Carriers_by_Tailnum)
```

## Summary

There were 17 planes that flew for multiple airlines. Planes N146PQ, N153PQ, N176PQ, N181PQ, N197PQ, N200PQ, N228PQ, and N232PQ flew for airlines 9E and EV. Planes N933AT, N935AT, N977AT, N978AT, N979AT, N981AT, N989AT, N990AT, and N994AT flew for airlines FL and DL.

# Question 4

In the figure above, there is a missing relationship between weather and airports. What is the relationship and how should it appear in the diagram?

```{r}
weather
airports
```

## Summary

The missing relationship between weather and airports is the airport of origin from which the weather observation was made. In weather, the airport of origin is under the column "origin." In airports, the column "faa" refers to the FAA code of the airport. However, it appears that there are only three origins in weather, those being EWR, LGA, and JFK, while airports has 1,458 FAA codes. In the diagram, a connection should be made between weather "origin" and airports "faa."

# Question 5

## Part 1

In order to address our primary question, we want to prepare a tidy dataset (a single data frame). This will consist of information from the flights and weather data. The year, month, day, hour, and origin variables from weather provide almost enough information to give each observation a unique value. Create a new variable in the weather dataset by pasting together those 5 variables, then determine how many duplicated values there are and explain why (Hint: compare your new variable to the time_hour variable).

```{r}
weather$Unique_Weather <- paste(weather$year, weather$month, weather$day, weather$hour, weather$origin)

sum(duplicated(weather$Unique_Weather))

weather[duplicated(weather$Unique_Weather, fromLast = TRUE), c("time_hour", "Unique_Weather")]
```

### Summary

There are 3 duplicated values in the newly created "Unique_Weather" variable, occurring on November 3rd, 2013 at 1:00 AM at each airport. This is because this day was the end of Daylight Saving time, and clocks were moved back an hour, so two observations of 1:00 AM were recorded at each airport.

## Part 2

Now merge the flights data and the weather data so that each flight contains information about the weather at its departure airport at the time it was scheduled to take off. You’ll want to use the variables time_hour and origin in order to do this.

```{r}
Merged_Flights_Weather <- merge(flights, weather, by = c("time_hour", "origin"), all.x = TRUE)

Merged_Flights_Weather <- subset(Merged_Flights_Weather, select = - c(year.y, month.y, day.y, hour.y))

names(Merged_Flights_Weather)[names(Merged_Flights_Weather) == "year.x"] <- "year"
names(Merged_Flights_Weather)[names(Merged_Flights_Weather) == "month.x"] <- "month"
names(Merged_Flights_Weather)[names(Merged_Flights_Weather) == "day.x"] <- "day"
names(Merged_Flights_Weather)[names(Merged_Flights_Weather) == "hour.x"] <- "hour"
```

# Question 6

On this merged dataset, perform steps 2-5 of the EDA checklist presented in class. Remember the context provided by our primary question above.

EDA Checklist: 2. Check the size of the data. 3. Examine the variables and their types. 4. Look at the top and bottom of the data. 5. Visualize the distributions of key variables.

```{r}
# 2. Check the size of the data
dim(Merged_Flights_Weather)
nrow(Merged_Flights_Weather)
ncol(Merged_Flights_Weather)

# 3. Examine the variables and their types
str(Merged_Flights_Weather)

# 4. Look at the top and bottom of the data
head(Merged_Flights_Weather)
tail(Merged_Flights_Weather)

# 5. Visualize the distributions of key variables
barplot(table(Merged_Flights_Weather$origin), main = "Number of Flights by Origin Airport", xlab = "Origin Airport", ylab = "Number of Flights", ylim = c(0, 140000))
mean(is.na(Merged_Flights_Weather$origin))

hist(Merged_Flights_Weather$dep_delay, main = "Distribution of Departure Delays", xlab = "Departure Delay (minutes)", sub = "Negative Values Represent Earlier Depature")
summary(Merged_Flights_Weather$dep_delay)
mean(is.na(Merged_Flights_Weather$dep_delay))

hist(Merged_Flights_Weather$arr_delay, main = "Distribution of Arrival Delays", xlab = "Arrival Delay (minutes)", sub = "Negative Values Represent Earlier Arrival")
summary(Merged_Flights_Weather$arr_delay)
mean(is.na(Merged_Flights_Weather$arr_delay))

hist(Merged_Flights_Weather$temp, main = "Distribution of Temperatures at Origin Airports", xlab = "Temperature (Fahrenheit)")
summary(Merged_Flights_Weather$temp)
mean(is.na(Merged_Flights_Weather$temp))

hist(Merged_Flights_Weather$dewp, main = "Distribution of Dewpoint at Origin Airports", xlab = "Dew Point (Fahrenheit)")
summary(Merged_Flights_Weather$dewp)
mean(is.na(Merged_Flights_Weather$dewp))

hist(Merged_Flights_Weather$humid, main = "Distribution of Humidity at Origin Airports", xlab = 
"Humidity")
summary(Merged_Flights_Weather$humid)
mean(is.na(Merged_Flights_Weather$humid))

hist(Merged_Flights_Weather$wind_speed, main = "Distribution of Wind Speed at Origin Airports", xlab = "Wind Speed (mph)")
summary(Merged_Flights_Weather$wind_speed)
mean(is.na(Merged_Flights_Weather$wind_speed))

hist(Merged_Flights_Weather$wind_gust, main = "Distribution of Wind Gust at Origin Airports", xlab = "Wind Gust (mph)")
summary(Merged_Flights_Weather$wind_gust)
mean(is.na(Merged_Flights_Weather$wind_gust))

hist(Merged_Flights_Weather$precip, main = "Distribution of Precipitation at Origin Airports", xlab = "Precipitation (inches)")
summary(Merged_Flights_Weather$precip)
mean(is.na(Merged_Flights_Weather$precip))

hist(Merged_Flights_Weather$pressure, main = "Distribution of Atmospheric Pressure at Origin Airports", xlab = "Atmospheric Pressure (millibars)")
summary(Merged_Flights_Weather$pressure)
mean(is.na(Merged_Flights_Weather$pressure))

hist(Merged_Flights_Weather$visib, main = "Distribution of Visibility at Origin Airports", xlab = "Visibility (miles)")
summary(Merged_Flights_Weather$visib)
mean(is.na(Merged_Flights_Weather$visib))
```

## Summary

There are 328,063 recordings (rows) and 32 variables (columns). The variables are mainly integer and numeric, but there is one logical variable and some character variables. In the context of the primary question, the key variables would be: dep_delay, arr_delay, temp, dewp, humid, wind_speed, wind_gust, precip, pressure, and visib. Measurements units were found using the data library for flights and weather data. The data was not cleaned in this step, so there are still some missing values.

# Question 7

## Part 1

Calculate the average departure delay for each day (Hint: think about which variables you need in order to do this!). Which day had the worst average length of delay for departures?

```{r}
Daily_Delay <- Merged_Flights_Weather |>
  group_by(year, month, day) |>
  summarize(Average_Departure_Delay = mean(dep_delay, na.rm = TRUE), Amount_of_Flights = n()) |>
  arrange(desc(Average_Departure_Delay))

head(Daily_Delay)
```

### Summary

March 8th, 2013 was the worst day, as it had an average departure delay of \~84 minutes.

## Part 2

Now calculate the averages by day and origin. Which airport had the worst single day for delays and when was it?

```{r}
Daily_Delay_Airport <- Merged_Flights_Weather |>
  group_by(year, month, day, origin) |>
  summarize (Average_Depatrue_Delay = mean(dep_delay, na.rm = TRUE), Amount_of_Flights = n()) |>
  arrange(desc(Average_Depatrue_Delay))

head(Daily_Delay_Airport)
```

### Summary

LGA had the single worst day for delays on March 8th, 2013, with an average departure delay of \~106 minutes.

## Part 3

Now calculate the averages by hour and origin. Which airport had the worst single hour for delays (and when was it)?

```{r}
Hourly_Delay_Airport <- Merged_Flights_Weather |>
  group_by(time_hour, origin) |>
  summarize(Average_Depature_Delay = mean(dep_delay, na.rm = TRUE), Amount_of_Flights = n()) |>
  arrange(desc(Average_Depature_Delay))

head(Hourly_Delay_Airport)
```

### Summary

LGA had the worst single hour for delays on July 22nd, 2013 at 6 PM, with an average departure delay of \~310 minutes.

# Question 8

Compute the average arrival delay by destination, then add that information to the airports data frame. Make a map showing the spatial distribution of delays (you can make it however you like, but we recommend using either ggplot or leaflet). Use the color of the points to display the average delay for each airport.

```{r}
Arrival_Delay <- Merged_Flights_Weather |>
  group_by(dest) |>
  summarize(Average_Arrival_Delay = mean(arr_delay, na.rm = TRUE), Amount_of_Flights = sum(!is.na(arr_delay))) |>
  filter(Amount_of_Flights > 0)

airports <- airports |>
  left_join(Arrival_Delay, by = c("faa" = "dest"))

airports <- airports |>
  filter(!is.na(Average_Arrival_Delay))

US_Map <- map_data("state")
ggplot() +
  geom_polygon(data = US_Map, aes(x = long, y = lat, group = group), fill = "gray", color = "white") +
  geom_point(data=airports,  aes(x = lon, y = lat, color = Average_Arrival_Delay)) + 
  scale_color_gradient2(low = "darkblue", mid = "orange", high = "cyan", name = "Average Arrival Delay (Minutes)") +
  labs(title = "Average Arrival Delay by Arrival Airport", x = "Longitude", y = "Latitude")
```

# Question 9

Finally, create as many plots and summary statistics as needed to answer the primary question: Which weather phenomena have the most impact on flight delays?

```{r}
Weather_Variables <- c("dep_delay", "temp", "humid", "dewp", "wind_speed", "wind_gust", "precip", "pressure", "visib")

Weather_Delay <- Merged_Flights_Weather
Weather_Delay$Visibility_Category <- ifelse(is.na(Weather_Delay$visib), NA, ifelse(Weather_Delay$visib < 3, "Low Visibility", "High Visibility"))
Weather_Delay$Precipitation_Category <- ifelse(is.na(Weather_Delay$precip), NA, ifelse(Weather_Delay$precip > 0, "Precipitation", "No Precipitation"))
Weather_Delay$WindSpeed_Category <- ifelse(is.na(Weather_Delay$wind_speed), NA, ifelse(Weather_Delay$wind_speed >= 30, "High Wind", "Low Wind"))
Weather_Delay$WindGust_Category <- ifelse(is.na(Weather_Delay$wind_gust), NA, ifelse(Weather_Delay$wind_gust >= 30, "High Gust", "Low Gust"))
Weather_Delay$Pressure_Category <- ifelse(is.na(Weather_Delay$pressure), NA, ifelse(Weather_Delay$pressure >= 1015.92, "High Pressure", "Low Pressure"))
Weather_Delay$Dewpoint_Category <- ifelse(is.na(Weather_Delay$dewp), NA, ifelse(Weather_Delay$dewp >= 65, "High Dew Point", "Low Dew Point"))
Weather_Delay$Temperature_Category <- ifelse(is.na(Weather_Delay$temp), NA, ifelse(Weather_Delay$temp >= 32, "Not Freezing", "Freezing"))
Weather_Delay$Humidity_Category <- ifelse(is.na(Weather_Delay$humid), NA, ifelse(Weather_Delay$humid >= 70, "High Humidity", "Low Humidity"))

Correlations <- cor(Weather_Delay[Weather_Variables], use = "pairwise.complete.obs", method = "spearman")

Departure_Delay_Correlation <- Correlations[rownames(Correlations) != "dep_delay", "dep_delay"]
Departure_Delay_Correlation <- Departure_Delay_Correlation[order(Departure_Delay_Correlation, decreasing = TRUE)]
Departure_Delay_Correlation

barplot(Departure_Delay_Correlation, horiz = TRUE, las = 1, col = ifelse(Departure_Delay_Correlation >0, "red", "blue"), xlab = "Correlation with Departure Delay", main = "Weather and Departure Delays")
```

```{r}
Visibility_Delay <- Weather_Delay |>
  filter(!is.na(Visibility_Category),!is.na(dep_delay)) |>
  group_by(origin, Visibility_Category) |>
  summarize(Departure_Delay_Mean = mean(dep_delay), Number_of_Flights = n(),.groups = "drop")

ggplot(Visibility_Delay,aes(x = origin, y = Departure_Delay_Mean,fill = Visibility_Category)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("High Visibility" = "red", "Low Visibility" = "blue")) +
  labs(title = "Average Departure Delay by Visibility", x = "Departure Airport", y = "Average Departure Delay (minutes)", subtitle = "A higher average for Low Visibility indicates a longer departure delay", fill = "Visibility Category")

<<<<<<< HEAD
plot(Weather_Delay$visib, Weather_Delay$dep_delay, xlab = "Visibility (miles)", ylab = "Departure Delay (minutes)", main = "Departure Delay by Visibility", xlim = c(0,10))
=======

Delayed <- !is.na(Weather_Delay$dep_delay) & Weather_Delay$dep_delay > 0
Amount_Delayed <- sum(Delayed)

plot(Weather_Delay$visib, Weather_Delay$dep_delay, xlab = "Visibility (miles)", ylab = "Departure Delay (minutes)", main = "Departure Delay by Visbility", xlim = c(0,10))
abline(lm(dep_delay ~ visib, data = Weather_Delay), col = "red")
>>>>>>> a379e6e10c0be15eee42538ee40df0dad6a92ab7
```

```{r}
Precipitation_Delay <- Weather_Delay |>
  filter(!is.na(Precipitation_Category),!is.na(dep_delay)) |>
  group_by(origin, Precipitation_Category) |>
  summarize(Departure_Delay_Mean = mean(dep_delay), Number_of_Flights = n(),.groups = "drop")

ggplot(Precipitation_Delay,aes(x = origin, y = Departure_Delay_Mean,fill = Precipitation_Category)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c( "No Precipitation" = "blue", "Precipitation" = "red")) +
  labs(title = "Average Departure Delay by Precipitation", x = "Departure Airport", y = "Average Departure Delay (minutes)", subtitle = "A higher average for Precipitation indicates a longer departure delay", fill = "Precipitation Category")

complete_rows <- complete.cases(Weather_Delay$precip, Weather_Delay$dep_delay)
plot(Weather_Delay$precip, Weather_Delay$dep_delay, xlab = "Precipitaton (inches)", ylab = "Departure Delay (minutes)", main = "Departure Delay by Precipitation")
lines(lowess(Weather_Delay$precip[complete_rows], Weather_Delay$dep_delay[complete_rows]), col = "red")
```

```{r}
WindSpeed_Delay <- Weather_Delay |>
  filter(!is.na(WindSpeed_Category),!is.na(dep_delay)) |>
  group_by(origin, WindSpeed_Category) |>
  summarize(Departure_Delay_Mean = mean(dep_delay), Number_of_Flights = n(),.groups = "drop")

ggplot(WindSpeed_Delay,aes(x = origin, y = Departure_Delay_Mean,fill = WindSpeed_Category)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("High Wind" = "red", "Low Wind" = "blue")) +
  labs(title = "Average Departure Delay by Wind Speed", x = "Departure Airport", y = "Average Departure Delay (minutes)", subtitle = "A higher average for High Wind Speed indicates a longer departure delay", fill = "Wind Speed Category")

complete_rows <- complete.cases(Weather_Delay$wind_speed, Weather_Delay$dep_delay)
plot(Weather_Delay$wind_speed, Weather_Delay$dep_delay, xlab = "Wind Speed (mph)", ylab = "Departure Delay (minutes)", main = "Departure Delay by Wind Speed")
lines(lowess(Weather_Delay$wind_speed[complete_rows], Weather_Delay$dep_delay[complete_rows]), col = "red")

WindGust_Delay <- Weather_Delay |>
  filter(!is.na(WindGust_Category),!is.na(dep_delay)) |>
  group_by(origin, WindGust_Category) |>
  summarize(Departure_Delay_Mean = mean(dep_delay), Number_of_Flights = n(),.groups = "drop")

ggplot(WindGust_Delay,aes(x = origin, y = Departure_Delay_Mean,fill = WindGust_Category)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("High Gust" = "red", "Low Gust" = "blue")) +
  labs(title = "Average Departure Delay by Wind Gust", x = "Departure Airport", y = "Average Departure Delay (minutes)", subtitle = "A higher average for High Wind Gust indicates a longer departure delay", fill = "Wind Gust Category")

complete_rows <- complete.cases(Weather_Delay$wind_gust, Weather_Delay$dep_delay)
plot(Weather_Delay$wind_gust, Weather_Delay$dep_delay, xlab = "Wind Gust (mph)", ylab = "Departure Delay (minutes)", main = "Departure Delay by Wind Gust")
lines(lowess(Weather_Delay$wind_gust[complete_rows], Weather_Delay$dep_delay[complete_rows]), col = "red")
```

```{r}
Pressure_Delay <- Weather_Delay |>
  filter(!is.na(Pressure_Category),!is.na(dep_delay)) |>
  group_by(origin, Pressure_Category) |>
  summarize(Departure_Delay_Mean = mean(dep_delay), Number_of_Flights = n(),.groups = "drop")

ggplot(Pressure_Delay,aes(x = origin, y = Departure_Delay_Mean,fill = Pressure_Category)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("High Pressure" = "red", "Low Pressure" = "blue")) +
  labs(title = "Average Departure Delay by Atmospheric Pressure", x = "Departure Airport", y = "Average Departure Delay (minutes)", subtitle = "A higher average for Low Pressure indicates a longer departure delay", fill = "Pressure Category")

complete_rows <- complete.cases(Weather_Delay$pressure, Weather_Delay$dep_delay)
plot(Weather_Delay$pressure, Weather_Delay$dep_delay, xlab = "Pressure (millibars)", ylab = "Departure Delay (minutes)", main = "Departure Delay by Atmospheric Pressure")
lines(lowess(Weather_Delay$pressure[complete_rows], Weather_Delay$dep_delay[complete_rows]), col = "red")
```

```{r}
Dewp_Delay <- Weather_Delay |>
  filter(!is.na(Dewpoint_Category),!is.na(dep_delay)) |>
  group_by(origin, Dewpoint_Category) |>
  summarize(Departure_Delay_Mean = mean(dep_delay), Number_of_Flights = n(),.groups = "drop")

ggplot(Dewp_Delay,aes(x = origin, y = Departure_Delay_Mean,fill = Dewpoint_Category)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("High Dew Point" = "red", "Low Dew Point" = "blue")) +
  labs(title = "Average Departure Delay by Dew Point", x = "Departure Airport", y = "Average Departure Delay (minutes)", subtitle = "A higher average for High Dew Point indicates a longer departure delay", fill = "Dew Point Category")

complete_rows <- complete.cases(Weather_Delay$pressure, Weather_Delay$dep_delay)
plot(Weather_Delay$dewp, Weather_Delay$dep_delay, xlab = "Dew Point (Fahrenheit)", ylab = "Departure Delay (minutes)", main = "Departure Delay by Dew Point")
lines(lowess(Weather_Delay$dewp[complete_rows], Weather_Delay$dep_delay[complete_rows]), col = "red")

Temp_Delay <- Weather_Delay |>
  filter(!is.na(Temperature_Category),!is.na(dep_delay)) |>
  group_by(origin, Temperature_Category) |>
  summarize(Departure_Delay_Mean = mean(dep_delay), Number_of_Flights = n(),.groups = "drop")

ggplot(Temp_Delay,aes(x = origin, y = Departure_Delay_Mean,fill = Temperature_Category)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("Not Freezing" = "red", "Freezing" = "blue")) +
  labs(title = "Average Departure Delay by Temperature", x = "Departure Airport", y = "Average Departure Delay (minutes)", subtitle = "A higher average for Not Freezing Temperature indicates a longer departure delay", fill = "Temperature Category")

complete_rows <- complete.cases(Weather_Delay$pressure, Weather_Delay$dep_delay)
plot(Weather_Delay$temp, Weather_Delay$dep_delay, xlab = "Temperature (Fahrenheit)", ylab = "Departure Delay (minutes)", main = "Departure Delay by Temperature")
lines(lowess(Weather_Delay$temp[complete_rows], Weather_Delay$dep_delay[complete_rows]), col = "red")

Humid_Delay <- Weather_Delay |>
  filter(!is.na(Humidity_Category),!is.na(dep_delay)) |>
  group_by(origin, Humidity_Category) |>
  summarize(Departure_Delay_Mean = mean(dep_delay), Number_of_Flights = n(),.groups = "drop")

ggplot(Humid_Delay,aes(x = origin, y = Departure_Delay_Mean,fill = Humidity_Category)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("High Humidity" = "red", "Low Humidity" = "blue")) +
  labs(title = "Average Departure Delay by Humidity", x = "Departure Airport", y = "Average Departure Delay (minutes)", subtitle = "A higher average for High Humidity indicates a longer departure delay", fill = "Humidity Category")

complete_rows <- complete.cases(Weather_Delay$pressure, Weather_Delay$dep_delay)
plot(Weather_Delay$humid, Weather_Delay$dep_delay, xlab = "Humidity", ylab = "Departure Delay (minutes)", main = "Departure Delay by Humidity")
lines(lowess(Weather_Delay$humid[complete_rows], Weather_Delay$dep_delay[complete_rows]), col = "red")

```
```{r}
Storm_Delay <- Weather_Delay |>
  mutate(Precipitation_Condition = precip > 0, Low_Visibility_Condition = visib < 3, High_Wind_Condition = wind_speed >= 30, High_Gust_Condition = wind_gust >= 30, Low_Pressure_Condition = pressure < 1015.92, Storm_Score = rowSums(cbind(Precipitation_Condition, Low_Visibility_Condition, High_Wind_Condition, High_Gust_Condition, Low_Pressure_Condition), na.rm = TRUE), Storm_Category = ifelse(Storm_Score >=3, "Storm Conditions", "Non-Storm Conditions"))

Hourly_Storm_Delay <- Storm_Delay |>
  group_by(origin, time_hour, month, Storm_Category) |>
  summarize(Average_Departure_Delay = mean(dep_delay, na.rm =TRUE), Amount_of_Flights = sum(!is.na(dep_delay))) |>
  filter(Amount_of_Flights > 0, is.finite(Average_Departure_Delay))

Monthly_Storm_Delay <- Hourly_Storm_Delay |>
  group_by(month, Storm_Category) |>
  summarize(Amount_of_Hours = n(), Amount_of_Flights = sum(Amount_of_Flights), Average_Departure_Delay = mean(Average_Departure_Delay, Amount_of_Flights)) |>
  mutate(Month = factor(month, levels = 1:12, labels = month.abb))
Monthly_Storm_Delay

ggplot(Monthly_Storm_Delay, aes(x = Month, y = Average_Departure_Delay, color = Storm_Category, group = Storm_Category)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  scale_color_manual(values = c("Storm Conditions" = "red", "Non-Storm Conditions" = "blue")) +
  labs(title = "Monthly Departure Delays by Weather Conditions", x = "Month", y = "Average Departure Delay (minutes)", color = "Weather Conditions")

ggplot(Hourly_Storm_Delay, aes(x = factor(month, levels = 1:12, labels = month.abb), y = Average_Departure_Delay, color = Storm_Category, group = Storm_Category)) +
  stat_summary(fun = mean, geom = "line", linewidth = 1) +
  stat_summary(fun = mean, geom = "point", size = 2.5) +
  facet_wrap(~ origin, ncol = 1) +
  scale_color_manual(values = c("Storm Conditions" = "red", "Non-Storm Conditions" = "blue")) +
  labs(title = "Monthly Departure Delays by Airport", x = "Month", y = "Average Departure Delay (minutes)", color = "Weather Condition")

```

## Notes

Low visibility was determined as anything less than 3 miles. High wind and high gust were determined to be anything 30+ mph. High pressure was determined to be anything above 1015.92 millibars (30 inHg). Ground de-icing occurs at 32°F, per FAA guidelines, and winter storms form at or below 32°F, per NOAA. High dew point was determined as anything greater than 60°F. High humidity was determined as anything greater than 70%. Lastly, precipitation was determined as anything greater than 0 inches of precipitation. https://www.faa.gov/nextgen/programs/weather/faq, https://www.faa.gov/air_traffic/publications/atpubs/aim_html/chap_7.html, https://www.faa.gov/other_visit/aviation_industry/airline_operators/airline_safety/deicing, https://www.nssl.noaa.gov/education/svrwx101/winter/, https://www.weather.gov/lmk/humidity, https://www.weather.gov/arx/heat_index, and https://www.weather.gov/arx/why_dewpoint_vs_humidity were used as sources for these metrics.

## Summary

Atmospheric pressure and visibility had weak negative associations with departure delay. Though flights experienced longer delays under low visibility and atmospheric pressure conditions, both of which accompany rain, snow, and storms, the correlations suggest that these two variables alone had limited influence on departure delays. Wind gust, temperature, wind speed, dew point, humidity, and precipitation had weak positive associations with departure delay, with precipitation having the strongest association at ~0.12. Flights experienced delays in high wind speed, wind gust, dew point, humidity, and "not freezing" temperature conditions, as well as when it was precipitating. These variables, besides "not freezing" temperatures, also accompany rain, snow, and storms, suggesting that winter storm conditions may have an impact on flight delays. However, the calculated correlations are weak, so weather alone cannot establish causation, but does provide some explanation for the delays. Along with that, "not freezing" temperatures were established as anything above 32°F, which is when water freezes. This means that temperatures just above 32°F are categorized as "not freezing". Dew point, humidity, and temperature interact with each other, which could make the air feel cooler than the raw temperature measurement. Again, all of these variables must be taken into account altogether, as their correlations were weak. 

This interaction made me want to find which months experienced the most "storm conditions." If an observation had three of the five variables that are associated with winter storms (wind speed, wind gust, dew point, and humidity), it was classified as having "storm conditions." The "Monthly Departure Delays by Weather Conditions" plot reveals that January, February, March, April, May, June, November, and December had higher average departure delays during storm conditions than non-storm condition observations. Given that winter lasts from December to March, it makes sense to see January, February, March, and December in this list. Given that spring lasts from March to May, that provides some explanation for April and May, as rain is common. June occurs during the summer, which may bring thunderstorms because of warmer temperatures and increased humidity. 
