---
title: "Lab 4"
author: "Nathan Weitershausen"
format: html
editor: visual
embed_resoures: true
---

# Step 1

```{r}
if (!file.exists("met_all.gz"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
    destfile = "met_all.gz",
    method   = "libcurl",
    timeout  = 60
    )
met <- read.csv("met_all.gz")
```

# Step 2

```{r}
met <- met[met$temp > -17, ]
met2 <- met[order(met$temp), ]
head(met2)

met2[met2==9999.0] <- NA

met2$date <- as.Date(paste(met2$year, met$month, met$day, sep= '-'))
summary(met2$date)

met2 <- met2[order(met2$date), ]
head(met2)

Met_FirstWeek <- met2[met2$day >= 1 & met2$day <= 7,]
Met_FirstWeek <- Met_FirstWeek[order(Met_FirstWeek$day),]

Station_Means <- aggregate(Met_FirstWeek[c("temp","rh","wind.sp","vis.dist","dew.point","lat","lon","elev")],
  by = list(station = Met_FirstWeek$USAFID),
  FUN = mean,
  na.rm= TRUE)

Station_Means$region <- ifelse(Station_Means$lat >= 39.71 & Station_Means$lon <= -98.00, "NW",
                               ifelse(Station_Means$lat < 39.71 & Station_Means$lon <= -98.00, "SW",
                                      ifelse(Station_Means$lat >= 39.71 & Station_Means$lon > -98.00, "NE", "SE")))

Station_Means$Elev_Cat <- ifelse(Station_Means$elev > 252, "high", "low")
```

# Step 3

```{r}
library(ggplot2)

Station_Means[!is.na(Station_Means$region),] |>
  ggplot() +
  geom_violin(mapping=aes(x=1, y=rh,fill=region))

Station_Means[!is.na(Station_Means$region),] |>
  ggplot() +
  geom_violin(mapping=aes(x=1, y=dew.point,fill=region))
```

Based on the graphs, it looks like dew point and relative humidity may be correlated, with regions with high relative humidity having high dew points and vice versa.

# Step 4

```{r}
Station_Means[!is.na(Station_Means$region), ] |>
  ggplot(mapping = aes(x = dew.point, y = rh, color = region)) + 
  geom_point(na.rm = TRUE)+
  geom_smooth(method = lm, se = FALSE, col="black")
```

Based on this graph, I see the relationship I described in the last step. Dew point and relative humidity are correlated.

# Step 5

```{r}
Station_Means[!is.na(Station_Means$Elev_Cat), ] |>
  ggplot() + 
  geom_bar(mapping = aes(x = Elev_Cat, fill = region), position = "dodge", na.rm = TRUE) + scale_fill_brewer (palette = "Set1") + labs(title = "Weather Stations by Elevation and Region Category", x = "Elevation Category", y = "Number of Weather Stations", fill = "Region")
```

According to the graph, the number of weather stations in the Northeast is almost evenly split between the "high" and "low" elevation categories. The number of weather stations in the Northwest are mainly in the "high" elevation category, while the opposite is true for the stations in the Southeast. The number of weather stations in the Southwest in the "high" elevation category appear to double that of the "low" elevation category.

# Step 6

```{r}
DP <- Station_Means[!is.na(Station_Means$dew.point), ] |>
  ggplot() + 
    stat_summary(mapping = aes(x = region, y = dew.point,),
    fun.data="mean_sdl",
    geom = "errorbar")
DP

WS <- Station_Means[!is.na(Station_Means$wind.sp), ] |>
  ggplot() + 
    stat_summary(mapping = aes(x = region, y = wind.sp,),
    fun.data="mean_sdl",
    geom = "errorbar")
WS
```

Dew point is higher in the Eastern regions than the Western regions. The Southeast has the highest mean dew point, followed by the Northeast, while the Northwest and Southwest have low mean dew points. The Eastern regions have smaller error bars, suggesting that they experience less variability. The Western regions have large error bars, suggesting that they experience a great amount of variability.

Wind speed is higher in the Western regions than the Eastern regions. The Southwest has the highest mean wind speed, followed closely by the Northwest. All regions have overlapping error bars, suggesting a lot of variability in each region.

# Step 7

```{r}
library(leaflet)
library(dplyr)
Station_Means <- summarize(Station_Means,
                      rh =     mean(rh, na.rm = TRUE),
                      lat =      mean(lat, na.rm = TRUE),
                      lon =      mean(lon, na.rm = TRUE),
                      .by = c(station))
Station_Means <- Station_Means[!is.na(Station_Means$rh), ]

rh.pal <- colorNumeric(c('darkgreen','goldenrod','brown'), domain=Station_Means$rh)

Station_Means <- Station_Means |>
  filter(!is.na(rh), !is.na(lat), !is.na(lon))

rh_map <- leaflet(Station_Means) |>
  addTiles() |>
  addCircles(lat = ~lat, lng = ~lon, label = ~paste0(round(rh,2), 'C'), color = ~ rh.pal(rh), opacity = 1, fillOpacity = 1, radius = 500) |>
  addLegend('bottomleft', pal=rh.pal, values=Station_Means$rh, title = 'Relative Humidity, %', opacity = 1) |>
  addMarkers(data = Station_Means |> filter(rank(-rh) <= 10))
  
rh_map
```

Relative humidity is relatively high in the Eastern portion of the U.S. Relative humidity begins to lower in the central and Southwestern U.S. before picking up again along the Western coastlines.

# Step 8

I selected 'gganimate' to animate the average temperature across all stations each day of the month.

```{r}
library(ggplot2)
library(gganimate)
library(gifski)

met2 <- met2[!is.na(met2$day), ]
met2$day <- as.numeric(met2$day)

avg <- ggplot(met2, aes(day, temp, group = 1)) +
  geom_line(color="red") +
  geom_point(color="red")+
  labs(title="Average Daily Temperature in August 2019", x = "Day", y = "Average Temperature") + 
  transition_reveal(day)

animation <- animate(avg, nframes = 50, fps = 10, width = 800, height = 500, units = "px", res = 50, renderer = gifski_renderer())

animation
```
