---
title: "Lab 6"
author: "Nathan Weitershausen"
format: html
editor: visual
embed-resources: true
---

```{r}
library(dplyr)
library(ggplot2)
library(tidytext)
library(readr)
library(tidyverse)
library(forcats)

mt_samples <- read_csv("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/00_mtsamples/mtsamples.csv")
mt_samples <- mt_samples |>
  select(description, medical_specialty, transcription)

head(mt_samples)
```

# Question 1: What specialties do we have?

Use the count() function from dplyr to figure out how many different categories we have in the data. Are these categories related? Overlapping? Evenly distributed?

```{r}
mt_samples |>
  count(medical_specialty, sort = TRUE)
```

## Summary
THere are 40 categories/medical specialties in the data. There doesn't seem to be any overalp and the categories are not evenly distributed. 

# Question 2
Tokenize the the words in the transcription column
Count the number of times each token appears
Visualize the top 20 most frequent words
Explain what we see from this result. Does it makes sense? What insights (if any) do we get?

```{r}
mt_samples |>
  unnest_tokens(token, transcription) |>
  count(token, sort = TRUE) |>
  top_n(20, n) |>
  ggplot(aes(n, fct_reorder(token,n))) + geom_col ()
```

## Summary
Based on the list, the top 20 most frequent words are stop words and "patient." This makes sense given stop words are common words and "patient" would be a common word given the context of the data. Because these are stop words, they do not give much context/insights, they actually carry little to no meaning at all.

# Question 3
Re-do the visualization but remove stop words before making it
Bonus points if you remove numbers as well
What do we see know that we have removed stop words? Does it give us a better idea of what the text is about?

```{r}
mt_samples |>
  unnest_tokens(token, transcription) |>
  anti_join(stop_words, by = c("token" = "word")) |>
  filter(!grepl("[0-9]", token)) |>
  count(token, sort=TRUE) |>
  top_n(20,n) |>
  ggplot(aes(n, fct_reorder(token,n))) + geom_col ()

```

## Summary
This list provides more context and reveals more "medical" terms, such as "procedure" and "diagnosis." This list makes more sense given the context of the data and appears to not have any leftover stopwords.

# Question 4
Repeat question 2, but this time tokenize into bi-grams. How does the result change if you look at tri-grams?
```{r}
mt_samples |>
  unnest_ngrams(ngram, transcription, n=2) |>
  count(ngram, sort= TRUE) |>
  top_n(20, n) |>
  ggplot(aes(n, fct_reorder(ngram,n))) + geom_col ()

mt_samples |>
  unnest_ngrams(ngram, transcription, n=3) |>
  count(ngram, sort= TRUE) |>
  top_n(20, n) |>
  ggplot(aes(n, fct_reorder(ngram,n))) + geom_col ()
```

## Summary
Tokenizing the data into bi-grams gives a list of mainly combined stopwords, for example, "of the" and "with a." Tokenizing the data into tri-grams gives a list that provides more context about the data, for example, "the operating room" and "past medical history." There is also a three-way tie for the 20th spot on the list, which is why there are 22 results rather than just 20. 

# Question 5
Using the results you got from Question 4, pick a word and count the words that appear before and after it.

```{r}
#Words that appear before "patient"
mt_samples |>
  unnest_ngrams(ngram, transcription, n=2) |>
  separate(ngram, into = c("word1", "word2"), sep = " ") |>
  select(word1, word2) |>
  filter(word2 == "patient") |>
  count(word1, sort= TRUE)

#Words that appear after "patient"
mt_samples |>
  unnest_ngrams(ngram, transcription, n=2) |>
  separate(ngram, into = c("word1", "word2"), sep = " ") |>
  select(word1, word2) |>
  filter(word1 == "patient") |>
  count(word2, sort= TRUE)
```

## Summary
The first list represents the words that appear before "patient" and the second list represents the words that appear after "patient."

# Question 6
Which words are most used in each of the specialties? You can use group_by() and top_n() from dplyr to have the calculations be done within each specialty. Remember to remove stop words. What are the 5 most-used words for each specialty?

```{r}
mt_samples |>
  unnest_tokens(token, transcription) |>
  group_by(medical_specialty)|>
  anti_join(stop_words, by = c("token" = "word")) |>
  filter(!grepl("[0-9]", token)) |>
  count(token, sort=TRUE) |>
  top_n(5,n) |>
  arrange(medical_specialty, desc(n))
```

## Summary 
The top five most used words for each specialty are lsited in this table. Some lists contain ties, so that is why more than five words appear for some specialties and not others. 

# Question 7
I wanted to see how different specialties referred to participants/patients/subjects/people, as I believe that will show the nature of the article/description, providing even more context.

```{r}
people_terms <- "^(patient|participant|subject|patients|participants|subjects|volunteer|volunteers|client|clinets)$"

mt_samples |>
  unnest_tokens(token, transcription) |>
  group_by(medical_specialty)|>
  filter(!grepl("[0-9]", token), grepl(people_terms, token)) |>
  count(token, sort= TRUE) |>
  arrange(medical_specialty, desc(n))
```

## Summary
From what I see, "patient" or "patients" seem to be the most common way these articles/descriptions refer to people, which makes sense given the graph in Question 3. By just combing the table, there aren't really any other words that stand out to me, as most have less than 10 counts. However, I do notice that some specialties use a variety of words, like "Chiropractic" and "Psychiatry/Psychology," while some specialties only use "patient" or "patients," like "ENT -Otolaryngology" and "Urology." If I were to expand the list of "people_terms" to include more words that could describe people, then maybe there would be more variation in the words used. 