---
title: "9/23 Slides"
author: "Nathan Weitershausen"
format: html
editor: visual
embed-resources: true
---

```{r}
library(data.table)
library(dtplyr)
library(dplyr)
library(ggplot2)
library(tidyverse)
```

```{r}
# Where are we getting the data from
met_url <- "https://github.com/USCbiostats/data-science-data/raw/master/02_met/met_all.gz"

# Downloading the data to a tempfile (so it is destroyed afterwards)
# you can replace this with, for example, your own data:
tmp <- tempfile(pattern = "met", fileext = ".gz")
# tmp <- "met.gz"

# We should be downloading this, ONLY IF this was not downloaded already.
# otherwise is just a waste of time.
if (!file.exists(tmp)) {
  download.file(
    url      = met_url,
    destfile = tmp,
    # method   = "libcurl", timeout = 1000 (you may need this option)
  )
}
```

```{r}
dat <- read.csv(tmp)
head(dat)
```
```{r}
# Selecting columns (method 1)
head(dat[,c('USAFID','lat','lon')])

#select one column
dat$month

#Using dplyr::select verb to select columns (method 2)
#pipe dat into select function, then pipe that into head
dat |> select(USAFID, lat, lon) |> head()

# select only the relevant variables
dat <- dat |> select(USAFID, WBAN, year, month, day, 
         hour, min, lat, lon, elev, 
         wind.sp, temp, atm.press)
```

# Data filtering: Logical Conditions
Based on logical operations: condition 1 and/or condition 2 and/or ...
Need to be aware of ordering and grouping of and and or operators

Fundamental relational operators: <,>, <=, >=, ==, !=

# How many ways can you write a Xor operator?

```{r}
myxor <- function(x, y) {
  res <- logical(length(x))
  for (i in 1:length(x)) {
    res[i]  # do something with x[i] and y[i]
  }
  return(res)
}

myxor1 <- function(x,y) {(x & !y) | (!x & y)}
myxor2 <- function(x,y) {!((!x | y) & (x | !y))}
myxor3 <- function(x,y) {(x | y) & (!x | !y)}
myxor4 <- function(x,y) {!((!x & !y) | (x & y))}
cbind(
  ifelse(xor(test[,1], test[,2]), "true", "false"),
  ifelse(myxor1(test[,1], test[,2]), "true", "false"),
  ifelse(myxor2(test[,1], test[,2]), "true", "false"),
  ifelse(myxor3(test[,1], test[,2]), "true", "false"),
  ifelse(myxor4(test[,1], test[,2]), "true", "false")
)
```

# Filtering (Subsetting) the data
Say we need to select records according to some criteria. For example:

First day of the month, and
Above latitude 40, and
Elevation outside the range 500 and 1,000.

he logical expressions would be

(day == 1)
(lat > 40)
((elev < 500) | (elev > 1000))

```{r}
#base r method
tmp <- dat[dat$day == 1 &
             dat$lat > 40 &
             (dat$elev < 500 | dat$elev > 1000), ]
head(tmp)

#dplyr::filter() method
dat |>
  filter(day == 1, lat > 40, (elev < 500) | (elev > 1000)) |>
  head() 
```

1. How many records have a tempertaure within 18 and 25?
2. Slme records hav missing values, count how mnay have temp or NA
3. Plot a sample of 1,000 pairs (lat,lon) coordinates for (a) the stations with temp as NA and (b) those with data

```{r}
# Question 1
nrow(dat[dat$temp < 25 & dat$temp > 18, ])
# dat[temp %between% c(18, 25), .N] 
# dat |> filter(between(temp, 18, 25)) |> nrow()

# Question 2
nrow(dat[is.na(dat$temp), ])
# more succinct: sum(is.na(dat$temp))

# Question 3
# Drawing a sample
set.seed(123)
idx1 <- sample(which(is.na(dat$temp)), 1000)
idx2 <- sample(which(!is.na(dat$temp)), 1000)

# Visualizing the data
# make a map of the US, as we did last class
ggplot(map_data("state"), aes(x = long, y = lat)) +
  geom_map(aes(map_id = region), map = map_data("state"), col = "lightgrey", fill = "gray") +
  geom_jitter(
    data    = dat[c(idx1, idx2), ],
    mapping = aes(x = lon, y = lat, col = is.na(temp)),
    inherit.aes = FALSE, alpha = .5, cex = 2
    )
```

# Creating Variables: Data Types
Logical: bool true/false type
Strings: string of characters
Integer: numeric variable without decimal (discrete)
Double: numeric variable with decimals (continuous)

# Creating Variables: Special Data Types
Time: could be date, date + time, or a combination
Categorical: represent strata/levels of variables
Ordinal: similar to facor, but has ordering

# Variable Creation
Want to creat two variables, elev^2 and scaled version of wind.sp by standard deviation
```{r}
dat$elev2 <- dat$elev^2
dat$windsp_scaled <- dat$wind.sp / sd(dat$wind.sp, na.rm = TRUE)
```

#variable creation using dplyr::mutate()
```{r}
dat |>
  mutate(
    elev2         = elev ^ 2,
    windsp_scaled = wind.sp / sd(wind.sp,na.rm=TRUE)
  ) |>
  head()
```

# Want to scale multiple variables by their SD but didn't want to copy-paste the code several times
```{r}
# Listing the names
names  <- c("wind.sp", "temp", "atm.press")

for(var in names){
  dat[,paste0(var,'_scaled')] <- dat[,var] / sd(dat[,var], na.rm = TRUE)
}

#Using dplyr
names  <- c("wind.sp", "temp", "atm.press")
dat |>
  mutate(
    across(
      all_of(names),
      function(x) x/sd(x, na.rm = TRUE),
      .names = "{col}_scaled2"
      )
  ) |>
  # Just to print the last columns
  select(ends_with("_scaled2")) |>
  head(n = 4)
```

# Complex Variable Creation

for loops and sapply may be slow but are handy for creating variables that rely on complicated relationships between variables

Complex variable: wind.temp - will take on 4 different values based on temperature and wind speed: cool & still, cool & windy, warm & still, warm & windy. SPlit each variable based on median value.

Will run slow and not be added as a variable to the dataset
```{r}
# create the new variable one entry at a time
wind.temp <- sapply(1:nrow(dat), function(i){
  if(is.na(dat$temp[i]) | is.na(dat$wind.sp[i])){
    return(NA)
  }
  if(dat$temp[i] <= median(dat$temp, na.rm=TRUE)){
    if(dat$wind.sp[i] <= median(dat$wind.sp, na.rm=TRUE)){
      return('cool & still')
    }else{
      return('cool & windy')
    }
  }else{
    if(dat$wind.sp[i] <= median(dat$wind.sp, na.rm=TRUE)){
      return('warm & still')
    }else{
      return('warm & windy')
    }
  }
})
```

For Loop Code (no sapply), unlike sapply, an empty vector must be made first and then filled in with the values.
```{r}
# initialize a variable of all missing values
wind.temp <- rep(NA, nrow(dat))
# fill in the values one at a time
for(i in 1:nrow(dat)){
  if(is.na(dat$temp[i]) | is.na(dat$wind.sp[i])){
    return(NA)
  }else{
    if(dat$temp[i] <= median(dat$temp, na.rm=TRUE)){
      if(dat$wind.sp[i] <= median(dat$wind.sp, na.rm=TRUE)){
        wind.temp[i] <- 'cool & still'
      }else{
        wind.temp[i] <- 'cool & windy'
      }
    }else{
      if(dat$wind.sp[i] <= median(dat$wind.sp, na.rm=TRUE)){
        wind.temp[i] <- 'warm & still'
      }else{
        wind.temp[i] <- 'warm & windy'
      }
    }
  }
}
```

Vectorized approach that will run on a large datset. 
```{r}
# initialize a variable of all missing values
wind.temp <- rep(NA, nrow(dat))
# assign every possible value by subsetting
wind.temp[dat$temp <= median(dat$temp, na.rm=TRUE) & 
            dat$wind.sp <= median(dat$wind.sp, na.rm=TRUE)] <- 'cool & still'
wind.temp[dat$temp <= median(dat$temp, na.rm=TRUE) & 
            dat$wind.sp > median(dat$wind.sp, na.rm=TRUE)] <- 'cool & windy'
wind.temp[dat$temp > median(dat$temp, na.rm=TRUE) & 
            dat$wind.sp <= median(dat$wind.sp, na.rm=TRUE)] <- 'warm & still'
wind.temp[dat$temp > median(dat$temp, na.rm=TRUE) & 
            dat$wind.sp > median(dat$wind.sp, na.rm=TRUE)] <- 'warm & windy'

head(wind.temp)
```

Merging Data
While building the MET dataset, we dropped the state data. We can use the original stations dataset and merge it to the MET dataset. Cannot do it right away, we need to process the data first.

```{r}
stations <- fread("https://noaa-isd-pds.s3.amazonaws.com/isd-history.csv")
stations <- as.data.frame(stations)
stations$USAF <- as.integer(stations$USAF)

# Dealing with NAs and 999999
stations$USAF[stations$USAF == 999999] <- NA
stations$CTRY[stations$CTRY == ""] <- NA
stations$STATE[stations$STATE == ""] <- NA

# Selecting the three relevant columns, and keep unique records
stations <- unique(stations[, c('USAF', 'CTRY', 'STATE')])

# Dropping NAs
stations <- stations[!is.na(stations$USAF), ]

head(stations, n = 4)
```
```{r}
merge(
  # Data
  x     = dat,      
  y     = stations, 
  # List of variables to match
  by.x  = "USAFID",
  by.y  = "USAF", 
  # Which obs to keep?
  all.x = TRUE,      
  all.y = FALSE
  ) |> nrow()
#Adds more rows, the original dataset has 2377343, this means there is duplicated IDs. Fix it with:

stations <- stations[!duplicated(stations$USAF), ]
```

```{r}
dat <- merge(
  # Data
  x     = dat,      
  y     = stations, 
  # List of variables to match
  by.x  = "USAFID",
  by.y  = "USAF", 
  # Which obs to keep?
  all.x = TRUE,      
  all.y = FALSE
  )

head(dat[, c('USAFID', 'WBAN', 'STATE')], n = 4)
```

# Aggregating Data: Adding grouped variables
```{r}
dat <- dat |>
  group_by(STATE, year, month, day) |>
  mutate(
    temp_imp = coalesce(temp, mean(temp, na.rm = TRUE))
    )

# Preparing for plotting
# select single station, sort by date
plotdata <- dat[dat$USAFID == 720172, ]
plotdata <- plotdata[order(plotdata$year, plotdata$month, plotdata$day, plotdata$hour, plotdata$min), ]
# Generate an 'x' variable for time
plotdata$id <- 1:nrow(plotdata)


plot(plotdata$id, plotdata$temp, pch=16)
lines(plotdata$id, plotdata$temp_imp, col=2)

plotdata |>
  ggplot(aes(x = id)) +
  geom_point(aes(y = temp)) +
  geom_line(aes(y = temp_imp, colour = 'red'))

head(dat)

dat <- ungroup(dat)
head(dat)
```

# Aggregate dat: Summary Table
```{r}
dat |> 
  group_by(STATE) |>
  summarise(
    temp_avg      = mean(temp, na.rm=TRUE),
    wind.sp_avg   = mean(wind.sp, na.rm=TRUE),
    atm.press_avg = mean(atm.press, na.rm = TRUE)
  ) |> 
  arrange(STATE) |> 
  head(n = 4)
```

