---
title: "Text Mining"
author: "Nathan Weitershausen"
format: html
editor: visual
embed-resources: true
---

# Text Mining

tidytext: work seemlessly with ggplot2, dplyr, tidyr

```{r}
library(tidytext)
```

# Alice's Adventures in Wonderland

```{r}
library(tidyverse)
alice <- readRDS("C:/Users/nweit/Downloads/alice.rds")
alice
```
# Tokenizing
Turning text into smaller units. Basically, we'll split by spaces in between text.

```{r}
library(tidytext)
alice |>
  unnest_tokens(token, text)
```

# Words as a unit
## Using dplyr verbs: count
```{r}
library(dplyr)
alice |>
  unnest_tokens(token, text) |>
  count(token)

library(dplyr)
alice |>
  unnest_tokens(token, text) |>
  count(token, sort = TRUE)

library(dplyr)
alice |>
  unnest_tokens(token, text) |>
  count(chapter, token)

library(dplyr)
alice |>
  unnest_tokens(token, text) |>
  group_by(chapter) |>
  count(token) |>
  top_n(10, n)
```

## dplyr verbs and ggplot2
```{r}
library(dplyr)
library(ggplot2)
alice |>
  unnest_tokens(token, text) |>
  count(token) |>
  top_n(10, n) |>
  ggplot(aes(n, token)) +
  geom_col()

library(dplyr)
library(ggplot2)
library(forcats)
alice |>
  unnest_tokens(token, text) |>
  count(token) |>
  top_n(10, n) |>
  ggplot(aes(n, fct_reorder(token, n))) +
  geom_col()
```

# Stop Words
These are words that do appear a lot but do not add much context: "the," "and," "at," and "for"

```{r} 
# tidytext comes with a data.frame of stop words
stop_words

# stop word lexicons
table(stop_words$lexicon)

# snowball stopwords
stop_words |>
  filter(lexicon == "snowball") |>
  pull(word)

# Duplicated stopwords
sort(table(stop_words$word), decreasing = TRUE)

```

```{r}
# Removing stopwords
#We can use anti_join() to remove the tokens that also appear in stop_words data.frame
alice |>
  unnest_tokens(token, text) |>
  anti_join(stop_words, by = c("token" = "word")) |>
  count(token, sort = TRUE)

# Anti-join with same variable name
alice |>
  unnest_tokens(word, text) |>
  anti_join(stop_words, by = c("word")) |>
  count(word, sort = TRUE)

# Stop words removed
alice |>
  unnest_tokens(word, text) |>
  anti_join(stop_words, by = "word") |>
  count(word, sort = TRUE) |>
  top_n(10, n) |>
  ggplot(aes(n, fct_reorder(word, n))) +
  geom_col()
```

# Which words appear together?
ngrams are sets of n consecutive words and we can count these to see which words appear together most frequently

ngrams with n =1 are unigrams ("which," "words," "appear," "together")
ngrams with n=2 are bigrams ("which words," "appear together")
ngrams with n=3 are trigrams ("which words appear", "together")

```{r}
alice |>
  unnest_ngrams(ngram, text, n = 2)

alice |>
  unnest_ngrams(ngram, text, n = 2) |>
  count(ngram, sort = TRUE)

alice |>
  unnest_ngrams(ngram, text, n = 2) |>
  separate(ngram, into = c("word1", "word2"), sep = " ") |>
  select(word1, word2)

alice |>
  unnest_ngrams(ngram, text, n = 2) |>
  separate(ngram, into = c("word1", "word2"), sep = " ") |>
  select(word1, word2) |>
  filter(word1 == "alice")

# Most common first words appearing after "alice"
alice |>
  unnest_ngrams(ngram, text, n = 2) |>
  separate(ngram, into = c("word1", "word2"), sep = " ") |>
  select(word1, word2) |>
  filter(word1 == "alice") |>
  count(word2, sort = TRUE)

# Most common second words appearing after "alice"
alice |>
  unnest_ngrams(ngram, text, n = 2) |>
  separate(ngram, into = c("word1", "word2"), sep = " ") |>
  select(word1, word2) |>
  filter(word2 == "alice") |>
  count(word1, sort = TRUE)
```

# TF-IDF
TF: Term Frequency (the more frequent a words is, the more important it is, excluding stop words)
IDF: Inverse Document Frequency (how rare/specific is a word)
TF-IDF: product of TF and IDF
TF gives weight to terms that appear a lot, IDF gives weight to terms that appears in a few documents

```{r}
# TF-IDF with tidytext
alice |>
  unnest_tokens(text, text) |>
  count(text, chapter) |>
  bind_tf_idf(text, chapter, n) |>
  arrange(desc(tf_idf))
```

# Sentiment Analysis
AKA "Opinion Mining." Isa way in which we can use computers to attempt to understand the feelings conveyed in a text. This relies on a large, human-compiled database of words with known associations or feelings ("positive," "negative," "joy," "surpirse," "disgust," etc)

```{r}
# Sentiment Lexicons- bing - contains a large list of words and a binary association ("positive," "negative")
library(textdata)
get_sentiments('bing')
# Sentiment Lexicons - afinn - assigns words a value between -5 and 5 that represents positivity or negativity
get_sentiments('afinn')
# Sentiment Lexicons - nrc - assigns each word an associated sentiment. Some words appear more than others because they have multiple associations
get_sentiments('nrc')

```

```{r}
# Use bing to assign "positive" and "negative" labels to as many words as possible in the book
alice |>
  unnest_tokens(word, text) |>
  inner_join(get_sentiments("bing"))

# Sentiment by chapter, do this by counting the number of "positive" words and subtracting "negative" words
diff_by_chap <- alice |>
  unnest_tokens(word, text) |>
  inner_join(get_sentiments("bing")) |> 
  group_by(chapter) |> 
  summarise(sentiment = sum(sentiment == "positive") - sum(sentiment == "negative"))

diff_by_chap

barplot(diff_by_chap$sentiment, names.arg = diff_by_chap$chapter)
```

```{r}
# Alternatively, we can use the afinn lexicon and quantify the "sentiment" of each chapter by the average of all words with numeric associations

avg_by_chap <- alice |>
  unnest_tokens(word, text) |>
  inner_join(get_sentiments("afinn")) |> 
  group_by(chapter) |> 
  summarise(sentiment = mean(value))

barplot(avg_by_chap$sentiment, names.arg = avg_by_chap$chapter)
```

```{r}
# Find the most frequent sentiment association in the nrc lexicon for each chapter. Unfortunately, the most frequent sentiment association ends up being "positive" or "negative"
alice |>
  unnest_tokens(word, text) |>
  inner_join(get_sentiments("nrc")) |> 
  group_by(chapter) |> 
  summarise(sentiment = names(which.max(table(sentiment))))


# Remove "positive" and "negative" from nrc lexicon
nrc_fun <- get_sentiments("nrc")
nrc_fun <- nrc_fun[!nrc_fun$sentiment %in% c("positive","negative"), ]
alice |>
  unnest_tokens(word, text) |>
  inner_join(nrc_fun) |> 
  group_by(chapter) |> 
  summarise(sentiment = names(which.max(table(sentiment))))
```

